---
title: "zementisr Quickstart Guide"
author: "Alexander Lemm"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zementisr Quickstart Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Authentication 

Zementis Server's REST API uses HTTP Basic Authentication. For each request the client needs to provide username and password. 

Since typing your password in the console is a bit too dangerous (you might accidentally share the `.Rhistory` file) and asking each time gets too cumbersome quickly, the zementisr package requires that you store your secrets and the base URL of your Zementis Server as environment variables in the `.Renviron` file in your home directory.

Please, make sure to set the following environment variables accordingly in your `.Renviron` file before using functions from the zementisr package.:

``` {r eval=FALSE}
ZEMENTIS_base_url = "http://localhost:9083"
ZEMENTIS_usr = "guybrush.threepwood"
ZEMENTIS_pwd = "bigwhoop"

```

## The `...` argument

Each function from the zementisr package comes with a `...` (dot-dot-dot) argument. It is used to pass on additional arguments to the underlying HTTP method from the `httr` package. This might be necessary if you need to set some curl options explicitly  via `httr::config()`.
